---
title: "Regression Attempts"
author: "Jordan Gropper"
date: '2022-08-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(fixest)
library(ggplot2)
library(tidyverse)
library(vtable)
library(Ecdat)
library(ggstance)
library(multcomp)
library(NHANES)
library(corrplot)
library(patchwork)
library(haven)
library(marginaleffects)
```


```{r}
clean_df<- read_csv('../RawData/final_clean.csv')
```





## 1 How has COVID affected the health of the retail industry, as measured by employment?
-Andrew:
```{r}


```


## 2 How has retail fared relative to other industries?
-Jordan:

```{r}
ques2 <- clean_df %>% dplyr::select(YEAR, MONTH, monthyear, CPSID, WTFINL, AGE, SEX, RACE, CLASSWKR, WHYUNEMP, WKSTAT, INDNAME, covid1_dummy, covid2_dummy, covid3_dummy, REGION)

#Should I control for region and for covid dummies?

ques2$employed_full <- ifelse(ques2$WKSTAT %in% c("Full-time hours (35+), usually full-time", "Full-time hours (35+), usually part-time for economic reasons", "Full-time hours (35+), usually part-time for non-economic reasons"), 1, 0)


ques2$job_loser <- ifelse(ques2$WHYUNEMP %in% c("Job loser - on layoff", "Other job loser"), 1, 0)

ques2$retail_or_not <- ifelse(ques2$INDNAME %in% c("Retail Trade"), 1, 0)

```



This first regression looks at how full time employment has faired over the span of time for the covid waves, and has controls for the different waves of covid, as well as region and implements the weights argument. 
```{r}


#mod1
ques2m1 <- feols(employed_full ~ INDNAME + YEAR + MONTH + REGION + covid1_dummy + covid2_dummy + covid3_dummy, data= ques2, 
                 weights = ques2$WTFINL)
etable(ques2m1) 


resid_1 <- resid(ques2m1)
plot(fitted(ques2m1), resid_1)


```


## WALD tests Question 2
```{r}
wald(ques2m1, c("YEAR", "MONTH")) #both are useful

wald(ques2m1, c("REGION")) # All regions are significant

wald(ques2m1, c( "covid1_dummy", "covid2_dummy", "covid3_dummy")) # all are significant and useful

```



Certainly seems like a binary variable logit model would be more appropriate. I want to verify how job loss looks compared to job maintained first before checking that. The variable Job loser helps look at if the respondent was layed off or lost their job in any way over the span of time.
```{r}
#mod2
ques2m2 <- feols(job_loser ~ retail_or_not + INDNAME + YEAR + MONTH | REGION + covid1_dummy + covid2_dummy + covid3_dummy, 
                 data = ques2, weights = ques2$WTFINL, vcov='hetero')
etable(ques2m2)

resid_2 <- resid(ques2m2)
plot(fitted(ques2m2), resid_2)

```

```{r}
wald(ques2m2, c("YEAR", "MONTH")) #both are useful

wald(ques2m2, c("REGION")) # All regions are significant

wald(ques2m2, c("covid1_dummy", "covid2_dummy", "covid3_dummy")) # all are significant and useful

```

There is a very minor difference between loss of job and those that became unemployed, so we observe that in this regression:
```{r}
#mod3
ques2$unemper <- ifelse(ques2$WKSTAT %in% c("Unemployed, seeking part-time work", "Unemployed, seeking full-time work"), 1, 0)

avg_unemper <- mean(ques2$unemper) #looking for what the probability of being generally unemployed is

ques2m3 <- feols(unemper ~ retail_or_not + YEAR + MONTH | REGION + covid1_dummy + covid2_dummy + covid3_dummy, 
                 data = ques2, weights = ques2$WTFINL, vcov='hetero')
etable(ques2m3)

resid_3 <- resid(ques2m3)
plot(fitted(ques2m3), resid_3)

```
```{r}
wald(ques2m3, c("YEAR", "MONTH")) #both are useful

wald(ques2m3, c("REGION")) # All regions are significant

wald(ques2m3, c("covid1_dummy", "covid2_dummy", "covid3_dummy")) # all are significant and useful

```


## Comparing Through Etable
```{r}
etable(ques2m1, ques2m2, ques2m3)
```



### Now a Logit response for question 2, which will more directly answer the question we are looking at: How did retail do as far as loss of jobs over time, compared with other industries
```{r}

ques2logit <- feglm(unemper ~ retail_or_not + YEAR + MONTH | REGION + covid1_dummy + covid2_dummy + covid3_dummy, 
                 data = ques2, weights = ques2$WTFINL, family = binomial(link = 'logit')) # iremoved the weights

log_m <- marginaleffects(ques2logit)
summary(log_m)


ques2 <- ques2 %>% mutate(logit_predict = predict(ques2logit, type="response", newdata=ques2))
ques2 <- ques2 %>% mutate(logit_index = predict(ques2logit, type="link", newdata=ques2))




ques2 %>% ggplot() +
  geom_line(aes(x=logit_index, y=logit_predict))
```

```{r}
ques2 %>% ggplot() +
  geom_line(aes(x=logit_index, y=logit_predict)) + geom_smooth(aes(x=logit_index, y=unemper))
```
# This shows that the lines are overlapping a good amount, meaning that our predictions are accurate.
#Now an attempt at making an interaction term:
```{r}
# Creating a dummy variable to see before and during covid. I will keep the assumption that March was the first wave of covid, but 
#the layoffs would have initiated after covid had lingered, so I am marking in April:

ques2$covid_april <- ifelse(ques2$monthyear >= '2020-04-01', 1, 0) # the 1 is layoffs in #april

ques2logit2 <- feglm(unemper ~ retail_or_not*covid_april + YEAR + MONTH | REGION + covid1_dummy + covid2_dummy + covid3_dummy, 
                 data = ques2, weights = ques2$WTFINL, family = binomial(link = 'logit')) # iremoved the weights

log_m2 <- marginaleffects(ques2logit2)
summary(log_m2)

etable(ques2logit2)
```
```{r}
ques2 <- ques2 %>% mutate(logit_predict2 = predict(ques2logit2, type="response", newdata=ques2))
ques2 <- ques2 %>% mutate(logit_index2 = predict(ques2logit2, type="link", newdata=ques2))



ques2 %>% ggplot() +
  geom_line(aes(x=logit_index2, y=logit_predict2)) + geom_smooth(aes(x=logit_index2, y=unemper))
```
# the interaction term enables interesting interpretation, but is a less accurate model. 



## 3 Retail needs to worry about who has money to spend - what has changed about who is working and earning money?##

-Brooke and Sean
```{r}



```





```{r}

sean_df_q3 <- clean_df %>% 
  mutate(Working = ifelse(clean_df$WKSTAT %in% c("Full-time hours (35+), usually full-time", "Full-time hours (35+), usually part-time for economic reasons", "Full-time hours (35+), usually part-time for non-economic reasons"), 1, 0) )


sean_df_q3$race_dummy <- as.factor(sean_df_q3$race_dummy)
sean_df_q3$student_dummy <- as.factor(sean_df_q3$student_dummy)



sean_df_q3$MARST %>% unique()
```


```{r}

#Tried turning employment into a binary variable and running a logit regression. I run into memory issues with the marginal effects so this won't work

# ques3m1 <- feglm(Working ~ CLASSWKR | covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, family = 'logit')
# 
# etable(ques3m1, se = 'hetero')
# 
# wald(ques3m1)
# 
# q3m1_me <- marginaleffects(ques3m1)
# summary(q3m1_me)
# 
# q3m1_pred <- sean_df_q3 %>% 
#   mutate(logit_predict = predict(log1), logit_index = predict(log1, type = 'link'))


```


```{r}

ques3m2 <- feols(IncNumber ~ AGE + race_dummy + student_dummy + MARST + YEAR |REGION + covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, vcov = 'hetero')

etable(ques3m2)

wald(ques3m2,"AGE")
wald(ques3m2,"race_dummy")
wald(ques3m2,"student_dummy")
wald(ques3m2,"MARST")
wald(ques3m2,"YEAR")

```
Odd that Age is slightly negative so lets take a look
```{r}

table(sean_df_q3$AGE, sean_df_q3$IncNumber)

plot <- sean_df_q3 %>%
  mutate(AGE = factor(AGE)) %>% 
  count(AGE, IncNumber) %>% 
  na.omit()

ggplot(plot, aes(x = AGE, y = n, fill = IncNumber)) +
  geom_col() +
  coord_flip()
```



```{r}

ques3m2.5 <- feols(IncNumber ~ AGE + I(AGE^2) | REGION + covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, vcov = 'hetero')

etable(ques3m2.5)

wald(ques3m2.5)
wald(ques3m2.5, "AGE")

```



```{r}

ques3m3 <- feols(IncNumber ~ race_dummy + student_dummy + MARST + YEAR | REGION + covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, vcov = 'hetero')

etable(ques3m3)

wald(ques3m3,"race_dummy")
wald(ques3m3,"student_dummy")
wald(ques3m3,"MARST")
wald(ques3m3,"YEAR")

```

```{r}

ques3m4 <- feols(IncNumber ~ AGE + I(AGE^2) + race_dummy + student_dummy + MARST + YEAR |REGION + covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, vcov = 'hetero')

etable(ques3m4)

wald(ques3m4,"AGE")
wald(ques3m4,"race_dummy")
wald(ques3m4,"student_dummy")
wald(ques3m4,"MARST")
wald(ques3m4,"YEAR")



```

The marriage variable being compared to is Divorced which makes sence. 

```{r}

etable(ques3m2,ques3m4,ques3m3)



```


```{r}
ques3m3 <- feols(IncNumber ~ race_dummy + student_dummy + MARST + YEAR | REGION + covid1_dummy + covid2_dummy + covid3_dummy  , data = sean_df_q3, weights = sean_df_q3$WTFINL, vcov = 'hetero')

etable(ques3m3)



```


































